---
title: "Impact_site_yld"
author: "Jackie Ouzman"
date: "24/09/2021"
output: html_document
---

```{r setup, include=FALSE}


library(ggplot2)
library(readxl)
library(tidyverse)

```

# Impact sites yield data.

This document is aimed to help work out how to group the trials for the impact site.

How to populate the economic analysis.

For example what yield values should I put in for year 1 to year 5 (we don't normally have this much data).

What are the likely costs and how will these change over time years 1 - 5.


## Multiple steps 

Data is sourced from what is on teams page and I am using the impact primary data here



```{r pre processing, include=FALSE}

trial_results <- read_csv("X:/Therese_Jackie/Sandy_soils/App_development2021/sandy_soil_tools_app/App_working/data/primary_data_all.csv")



```

## Plots of yield data for following sites

   
Tempy        
Wynarka      
Monia_Gap          
Malinong  
Sherwood  
Cummins    
Karkoo           
Buckleboo      
Mt_Damper          
Kybunga    
Warnertown      
Telopea_Downs  
Kooloonong_chickpea
Kooloonong_lentil  
Kooloonong_lupin

```{r Tempy working, include=FALSE}

site_selection <- "Tempy"
site <- trial_results %>% 
  filter(site == site_selection)

str(site)


### arrange the data so the order reflect level of intervension
unique(as.character(site$Descriptors))
unique(as.character(site$grouping))

order <- c("Control",
           "Rip.50_none",
           "Rip.50_Cl.deep", 
           "Rip.50IncRip_none", 
           "Rip.50IncRip_Cl.mix")

site$Descriptors <- factor(site$Descriptors, 
                            levels = order)
order_grouping <- c("control",
           "deep ripping no amendment",
           "deep riping amendment", 
           "inclusion ripping no amendment", 
           "inclusion ripping amendment")



site_name = unique(site$site)
year_amelioration = unique(site$Amelioration_Year)

decile_year0_temp <-filter(site, yr_post_amelioration == 0 ) %>% 
  select(decile)
decile_year1_temp <-filter(site, yr_post_amelioration == 1 ) %>% 
  select(decile)  

decile_year0 = as.character(unique(decile_year0_temp$decile))
decile_year1 = as.character(unique(decile_year1_temp$decile))


##########################################################################################################
#####################################     graph with   Descriptors   ##################################################

## I would like to highlight the control but I am having trouble getting this to work

# Add a column indicating whether the category should be highlighted
site <- site %>% mutate(Control = ifelse( Descriptors == "Control", "yes", NA ) )



site_plot_descriptors <- ggplot (site, aes(x = Descriptors)) +
  geom_bar(
    #aes(y = yield, fill = as.factor(yr_post_amelioration), color = Control), #this works but I can't remove the leg
    aes(y = yield, fill = as.factor(yr_post_amelioration)),
    stat = "summary",
    fun.y = "mean",
    position = position_dodge(0.8),
    width = 0.7,
    show.legend = TRUE
  ) +
 
  guides(fill = guide_legend(title = "Years post amelioration")) +
  scale_fill_grey() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  labs(title= paste0("Site: ", site_name, ", Year amelioration: ", year_amelioration),
       subtitle = paste0("Deciles: year 0 = " , decile_year0, " , year 1 = ", decile_year1),
       x ="Trial", y = "yield t/ha")


##########################################################################################################
#####################################     tables   with   Descriptors  ##################################################
## pull out some values to chcek graph and use to fill out the 5 years
Descriptors_table_site <- site %>% group_by(as.factor(yr_post_amelioration), Descriptors) %>% 
  summarise(yield_av = mean(yield , na.rm = TRUE)) %>% 
  rename(yr_post_amelioration= `as.factor(yr_post_amelioration)`)

Descriptors_table_site

Descriptors_table_site_display <- pivot_wider(Descriptors_table_site,
            names_from = yr_post_amelioration,
            values_from = yield_av)

Descriptors_table_site_display <-Descriptors_table_site_display %>%
  slice(match(order, Descriptors))
Descriptors_table_site_display


##########################################################################################################
#####################################     graph with   Descriptors   ##################################################

## I would like to highlight the control but I am having trouble getting this to work

# Add a column indicating whether the category should be highlighted
site <- site %>% mutate(Control = ifelse( Descriptors == "Control", "yes", NA ) )



site_plot_grouping <- ggplot (site, aes(x = grouping)) +
  geom_bar(
    #aes(y = yield, fill = as.factor(yr_post_amelioration), color = Control), #this works but I can't remove the leg
    aes(y = yield, fill = as.factor(yr_post_amelioration)),
    stat = "summary",
    fun.y = "mean",
    position = position_dodge(0.8),
    width = 0.7,
    show.legend = TRUE
  ) +
  
  guides(fill = guide_legend(title = "Years post amelioration")) +
  scale_fill_grey() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  labs(title= paste0("Site: ", site_name, ", Year amelioration: ", year_amelioration),
       subtitle = paste0("Deciles: year 0 = " , decile_year0, " , year 1 = ", decile_year1),
       x ="Trial", y = "yield t/ha")

site_plot_grouping

##########################################################################################################
#####################################     tables   with   grouping  ##################################################
## pull out some values to chcek graph and use to fill out the 5 years
grouping_table_site <- site %>% group_by(as.factor(yr_post_amelioration), grouping) %>% 
  summarise(yield_av = mean(yield , na.rm = TRUE)) %>% 
  rename(yr_post_amelioration= `as.factor(yr_post_amelioration)`)

grouping_table_site

grouping_table_site_display <- pivot_wider(grouping_table_site,
                                              names_from = yr_post_amelioration,
                                              values_from = yield_av)

grouping_table_site_display <-grouping_table_site_display %>%
  slice(match(order_grouping, grouping))
grouping_table_site_display




```

# Tempy plots and tables of yield data using *descriptors* to describe treatments

```{r Tempy descriptors , echo=FALSE}
Descriptors_table_site_display
site_plot_descriptors



```

### Tempy plots and tables of yield data using *aggregated* descriptors to describe treatments

```{r Tempy grouping , echo=FALSE}


grouping_table_site_display
site_plot_grouping

```


# Wynarka plots and tables of yield data using *descriptors* to describe treatments

```{r Wynarka working, include=FALSE}

site_selection <- "Wynarka"
site <- trial_results %>% 
  filter(site == site_selection)

str(site)


### arrange the data so the order reflect level of intervension
unique(as.character(site$Descriptors))
unique(as.character(site$grouping))

order <- c("Control",
           "Rip.30_none",
           "Rip.60_none",
           "Rip.60IncRip_none",
           "Rip.60IncRip+Spade.30_none",
           "Spade.30_none" ,              
           "Rip.60Spade.30_none")

site$Descriptors <- factor(site$Descriptors, 
                           levels = order)
order_grouping <- c("control" ,
                    "deep ripping no amendment",
                    "inclusion ripping no amendment" ,
                    "inclusion ripping + spading no amendment",
                    "rotary spading no amendment")


site_name = unique(site$site)
year_amelioration = unique(site$Amelioration_Year)

decile_year0_temp <-filter(site, yr_post_amelioration == 0 ) %>% 
  select(decile)
decile_year1_temp <-filter(site, yr_post_amelioration == 1 ) %>% 
  select(decile)  

decile_year0 = as.character(unique(decile_year0_temp$decile))
decile_year1 = as.character(unique(decile_year1_temp$decile))


##########################################################################################################
#####################################     graph with   Descriptors   ##################################################

## I would like to highlight the control but I am having trouble getting this to work

# Add a column indicating whether the category should be highlighted
site <- site %>% mutate(Control = ifelse( Descriptors == "Control", "yes", NA ) )



site_plot_descriptors <- ggplot (site, aes(x = Descriptors)) +
  geom_bar(
    #aes(y = yield, fill = as.factor(yr_post_amelioration), color = Control), #this works but I can't remove the leg
    aes(y = yield, fill = as.factor(yr_post_amelioration)),
    stat = "summary",
    fun.y = "mean",
    position = position_dodge(0.8),
    width = 0.7,
    show.legend = TRUE
  ) +
 
  guides(fill = guide_legend(title = "Years post amelioration")) +
  scale_fill_grey() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  labs(title= paste0("Site: ", site_name, ", Year amelioration: ", year_amelioration),
       subtitle = paste0("Deciles: year 0 = " , decile_year0, " , year 1 = ", decile_year1),
       x ="Trial", y = "yield t/ha")


##########################################################################################################
#####################################     tables   with   Descriptors  ##################################################
## pull out some values to chcek graph and use to fill out the 5 years
Descriptors_table_site <- site %>% group_by(as.factor(yr_post_amelioration), Descriptors) %>% 
  summarise(yield_av = mean(yield , na.rm = TRUE)) %>% 
  rename(yr_post_amelioration= `as.factor(yr_post_amelioration)`)

Descriptors_table_site

Descriptors_table_site_display <- pivot_wider(Descriptors_table_site,
            names_from = yr_post_amelioration,
            values_from = yield_av)

Descriptors_table_site_display <-Descriptors_table_site_display %>%
  slice(match(order, Descriptors))
Descriptors_table_site_display


##########################################################################################################
#####################################     graph with   Descriptors   ##################################################

## I would like to highlight the control but I am having trouble getting this to work

# Add a column indicating whether the category should be highlighted
site <- site %>% mutate(Control = ifelse( Descriptors == "Control", "yes", NA ) )



site_plot_grouping <- ggplot (site, aes(x = grouping)) +
  geom_bar(
    #aes(y = yield, fill = as.factor(yr_post_amelioration), color = Control), #this works but I can't remove the leg
    aes(y = yield, fill = as.factor(yr_post_amelioration)),
    stat = "summary",
    fun.y = "mean",
    position = position_dodge(0.8),
    width = 0.7,
    show.legend = TRUE
  ) +
  
  guides(fill = guide_legend(title = "Years post amelioration")) +
  scale_fill_grey() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  labs(title= paste0("Site: ", site_name, ", Year amelioration: ", year_amelioration),
       subtitle = paste0("Deciles: year 0 = " , decile_year0, " , year 1 = ", decile_year1),
       x ="Trial", y = "yield t/ha")

site_plot_grouping

##########################################################################################################
#####################################     tables   with   grouping  ##################################################
## pull out some values to chcek graph and use to fill out the 5 years
grouping_table_site <- site %>% group_by(as.factor(yr_post_amelioration), grouping) %>% 
  summarise(yield_av = mean(yield , na.rm = TRUE)) %>% 
  rename(yr_post_amelioration= `as.factor(yr_post_amelioration)`)

grouping_table_site

grouping_table_site_display <- pivot_wider(grouping_table_site,
                                              names_from = yr_post_amelioration,
                                              values_from = yield_av)

grouping_table_site_display <-grouping_table_site_display %>%
  slice(match(order_grouping, grouping))
grouping_table_site_display




```



```{r Wynarka descriptors , echo=FALSE}
Descriptors_table_site_display
site_plot_descriptors



```

Wynarka plots and tables of yield data using *aggregated* descriptors to describe treatments

```{r Wynarka grouping , echo=FALSE}


grouping_table_site_display
site_plot_grouping

```


# Monia_Gap plots and tables of yield data using *descriptors* to describe treatments

```{r Monia_Gap working, include=FALSE}

site_selection <- "Monia_Gap"
site <- trial_results %>% 
  filter(site == site_selection)

str(site)


### arrange the data so the order reflect level of intervension
unique(as.character(site$Descriptors))
unique(as.character(site$grouping))



order <- c("Control",
           "Rip.60Spade.30_none",
           "Spade.30_none",
           "Spade.30_gypsum.mix",  
           "Spade.30_Cl.mix",
           "Spade.30_Cl.gypsum.mix")
           

site$Descriptors <- factor(site$Descriptors, 
                           levels = order)

  
order_grouping <- c("control" ,
                    "inclusion ripping + spading no amendment",
                    "rotary spading amendment" ,
                    "rotary spading no amendment"   )


site_name = unique(site$site)
year_amelioration = unique(site$Amelioration_Year)

decile_year0_temp <-filter(site, yr_post_amelioration == 0 ) %>% 
  select(decile)
decile_year1_temp <-filter(site, yr_post_amelioration == 1 ) %>% 
  select(decile)  

decile_year0 = as.character(unique(decile_year0_temp$decile))
decile_year1 = as.character(unique(decile_year1_temp$decile))


##########################################################################################################
#####################################     graph with   Descriptors   ##################################################

## I would like to highlight the control but I am having trouble getting this to work

# Add a column indicating whether the category should be highlighted
site <- site %>% mutate(Control = ifelse( Descriptors == "Control", "yes", NA ) )



site_plot_descriptors <- ggplot (site, aes(x = Descriptors)) +
  geom_bar(
    #aes(y = yield, fill = as.factor(yr_post_amelioration), color = Control), #this works but I can't remove the leg
    aes(y = yield, fill = as.factor(yr_post_amelioration)),
    stat = "summary",
    fun.y = "mean",
    position = position_dodge(0.8),
    width = 0.7,
    show.legend = TRUE
  ) +
 
  guides(fill = guide_legend(title = "Years post amelioration")) +
  scale_fill_grey() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  labs(title= paste0("Site: ", site_name, ", Year amelioration: ", year_amelioration),
       subtitle = paste0("Deciles: year 0 = " , decile_year0, " , year 1 = ", decile_year1),
       x ="Trial", y = "yield t/ha")


##########################################################################################################
#####################################     tables   with   Descriptors  ##################################################
## pull out some values to chcek graph and use to fill out the 5 years
Descriptors_table_site <- site %>% group_by(as.factor(yr_post_amelioration), Descriptors) %>% 
  summarise(yield_av = mean(yield , na.rm = TRUE)) %>% 
  rename(yr_post_amelioration= `as.factor(yr_post_amelioration)`)

Descriptors_table_site

Descriptors_table_site_display <- pivot_wider(Descriptors_table_site,
            names_from = yr_post_amelioration,
            values_from = yield_av)

Descriptors_table_site_display <-Descriptors_table_site_display %>%
  slice(match(order, Descriptors))
Descriptors_table_site_display


##########################################################################################################
#####################################     graph with   Descriptors   ##################################################

## I would like to highlight the control but I am having trouble getting this to work

# Add a column indicating whether the category should be highlighted
site <- site %>% mutate(Control = ifelse( Descriptors == "Control", "yes", NA ) )



site_plot_grouping <- ggplot (site, aes(x = grouping)) +
  geom_bar(
    #aes(y = yield, fill = as.factor(yr_post_amelioration), color = Control), #this works but I can't remove the leg
    aes(y = yield, fill = as.factor(yr_post_amelioration)),
    stat = "summary",
    fun.y = "mean",
    position = position_dodge(0.8),
    width = 0.7,
    show.legend = TRUE
  ) +
  
  guides(fill = guide_legend(title = "Years post amelioration")) +
  scale_fill_grey() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  labs(title= paste0("Site: ", site_name, ", Year amelioration: ", year_amelioration),
       subtitle = paste0("Deciles: year 0 = " , decile_year0, " , year 1 = ", decile_year1),
       x ="Trial", y = "yield t/ha")

site_plot_grouping

##########################################################################################################
#####################################     tables   with   grouping  ##################################################
## pull out some values to chcek graph and use to fill out the 5 years
grouping_table_site <- site %>% group_by(as.factor(yr_post_amelioration), grouping) %>% 
  summarise(yield_av = mean(yield , na.rm = TRUE)) %>% 
  rename(yr_post_amelioration= `as.factor(yr_post_amelioration)`)

grouping_table_site

grouping_table_site_display <- pivot_wider(grouping_table_site,
                                              names_from = yr_post_amelioration,
                                              values_from = yield_av)

grouping_table_site_display <-grouping_table_site_display %>%
  slice(match(order_grouping, grouping))
grouping_table_site_display




```



```{r Monia_Gap descriptors , echo=FALSE}
Descriptors_table_site_display
site_plot_descriptors



```

Monia_Gap plots and tables of yield data using *aggregated* descriptors to describe treatments

```{r Wynarka grouping , echo=FALSE}



grouping_table_site_display
site_plot_grouping

```