---
title: "Control yields"
author: "Jackie Ouzman"
date: "10/02/2022"
output: html_document
---

```{r setup, include=FALSE}

library(ggplot2)
library(readxl)
library(tidyverse)
#library(multcompView)

library(plotly)

library(tidyverse)
library(hrbrthemes)
library(viridis)
library(forcats)
library(hrbrthemes)
#install.packages("hrbrthemes")


```

## Control yields

This is for a select number of sites:

##### Site in merged control files
•	Brimpton Lake

•	Brooker

•	Buckleboo

•	Bute_CSIRO

•	Bute_Trengrove

•	Cadgee

•	CarwarpAmelioration

•	Cummins

•	Karkoo

•	Karoonda

•	Kooloonong_canola

•	Kooloonong_chickpea

•	Kooloonong_lentil

•	Kooloonong_lupin

•	Lowaldie_Crest

•	Lowaldie_Deep sand

•	Mt_Damper

•	Murlong

•	Ouyen_spade

•	Telopea_Downs

•	Tempy

•	Waikerie

•	Wynarka

•	Yenda

•	Younghusband (only one year)

##### Not ready
•	Kybunga

•	Malinong

•	Monia_Gap

•	Sherwood

•	Warnertown






Note everything runs off of the rep specified in the raw data.
I have run into problems with Younghusband, because the rep are not like the other sites.
For the control the raw data has 4 yield results that are labelled rep 1.

This means that the ID (which is supposed to unique is not)

I have created a new rep which is the clm called bay number and the rep to create a rep_block
so for the control I have how have 
rep block called
Bay 1 rep 1
Bay 1 rep 2
....


```{r Get the data, include=FALSE}
#################################################################################################################
####   Get the data #####



#####################################################################################################################

data_file <- "X:/Therese_Jackie/Sandy_soils/Development_database/other_sites_working/stats_working/sites_merged.csv"
data_file_control <- "X:/Therese_Jackie/Sandy_soils/Development_database/other_sites_working/stats_working/sites_control_merged.csv"

## download the data using the specified file path above

summary_data_all <- read_csv(data_file, 
                             col_types = cols(rep_block = col_character()))
summary_control_data_all <- read_csv(data_file_control, 
                            col_types = cols(rep_block = col_character()))


unique(summary_control_data_all$site_sub)
unique(summary_control_data_all$site)

str(summary_control_data_all)



##label problem with Mt Damper
summary_control_data_all <- summary_control_data_all %>% 
  mutate(
    site_sub = case_when(
      site_sub == "Mt Damper" ~ "Mt_Damper",
      TRUE ~ site_sub))

##label problem with Ouyen
summary_control_data_all <- summary_control_data_all %>% 
  mutate(
    site_sub = case_when(
      site_sub == "Oyen_Spade" ~ "Ouyen_Spade",
      TRUE ~ site_sub))
summary_control_data_all <- summary_control_data_all %>% 
  mutate(
    site = case_when(
      site == "Oyen_Spade" ~ "Ouyen_Spade",
      TRUE ~ site))


## list of sites that are ready ish

list_sites_ready <- c("Brimpton Lake",
                      "Brooker",
                      "Buckleboo",
                      "Bute_CSIRO",
                      "Bute_Trengrove",
                      "Cadgee",
                      "Carwarp_Amelioration",
                      "Cummins",
                      "Karkoo",
                      "Karoonda",
                      #"Kooloonong_canola",
                      "Kooloonong_chickpea",
                      "Kooloonong_lentil",
                      "Kooloonong_lupin",
                      "Lowaldie_Crest",
                      "Lowaldie_Deep sand",
                      "Mt Damper",
                      "Mt_Damper",
                      "Murlong",
                      "Oyen_Spade",
                      "Ouyen_Spade",
                      "Telopea_Downs",
                      "Tempy",
                      "Waikerie",
                      "Wynarka",
                      "Yenda",
                      "Younghusband")


summary_control_data_all <- summary_control_data_all %>% 
  filter(site_sub %in% list_sites_ready  )
summary_data_all <- summary_data_all %>% 
  filter(site_sub %in% list_sites_ready  )





# write.csv(summary_control_data_all, 
# "X:/Therese_Jackie/Sandy_soils/Development_database/other_sites_working/stats_working/sites_merged_CONTROL_incomplete.csv",
# row.names = FALSE)
# 
# write.csv(summary_data_all, 
# "X:/Therese_Jackie/Sandy_soils/Development_database/other_sites_working/stats_working/sites_merged_incomplete.csv",
# row.names = FALSE)

```

How many sites?

```{r how many sites all, echo=FALSE, message=FALSE, warning=FALSE}
#How many site?

sites <- summary_data_all %>% 
  distinct(site) 
sites %>% 
  count()
```


How many desciptors?

```{r how many desciptors all, echo=FALSE, message=FALSE, warning=FALSE}
## how many desciptors?

#names(summary_data_all)
desciptors <- summary_data_all %>% 
  distinct(Descriptors) 
desciptors %>% 
  count()
```

How many soil modification (including different depths)?

```{r how many modification all, echo=FALSE, message=FALSE, warning=FALSE}
## how many different modification?

summary_data_all <- summary_data_all %>% 
  mutate(Descriptors_temp = Descriptors) %>% 
  separate(Descriptors_temp, c("modification_with_depth", "amendment"), sep = "_")

modification_with_depth <- summary_data_all %>% 
  distinct(modification_with_depth) %>% 
  arrange(modification_with_depth)
modification_with_depth %>% 
  count()


```

How many soil amendment (including different depths)?

```{r how many amendment all, echo=FALSE, message=FALSE, warning=FALSE}
amendment_with_depth <- summary_data_all %>% 
  distinct(amendment) %>% 
  arrange(amendment) %>% 
  filter(!is.na(amendment)) %>% 
  filter(amendment != "none" ) 


amendment_with_depth %>% 
  count()

```



## Scatter Plots



```{r plot all, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = summary_control_data_all, mapping = aes(control_yield, yield,)) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  geom_point(alpha= 0.4) +
  geom_smooth(method = lm, se = FALSE) +
  scale_x_continuous(breaks=seq(0,10,by=0.5))+
  scale_y_continuous(breaks=seq(0,10,by=0.5))+
  theme_bw()+
  labs(title = "Control yield - all data - no summary\nnote each site, treatment, year and rep is matched to control",
       x = "control yield t/ha", y = "trial yield t/ha")
```

```{r scatter plot all, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}

# every trial yield point that has a matching control yld is displayed

scatter_control_vs_trial <- ggplot(data = summary_control_data_all, mapping = aes(control_yield, yield,)) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  geom_point(alpha= 0.4) +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Control yield - all data - no summary\nnote each site, treatment, year and rep is matched to control",
       x = "control yield t/ha", y = "trial yield t/ha")

scatter_control_vs_trial

ggplotly(scatter_control_vs_trial)


```

## Scatter Plots with Facet wrap is years post amelioration, the NA related to trial with no metadata

Missing year trial start date for;


Kooloong Canola (missing I need to check what I have)

Ouygen Sapde (missing)

```{r scatter plot all facet years post, echo=FALSE, message=FALSE, warning=FALSE}


ggplot(data = summary_control_data_all, mapping = aes(control_yield, yield,)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  # scale_x_continuous(breaks=seq(0,10,by=0.5))+
  # scale_y_continuous(breaks=seq(0,10,by=0.5))+
  theme_bw()+
  labs(title = "Control yield - all data - no summary\nnote each site, treatment, year and rep is matched to control
       \nFacet wrap is years post amelioration, the NA related to trial with no metdata",
       x = "control yield t/ha", y = "trial yield t/ha")+
  facet_wrap(.~ yr_post_amelioration)


```


## Scatter Plots with Facet wrap is sites


```{r scatter plot all facet sub site post, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = summary_control_data_all, mapping = aes(control_yield, yield,)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  # scale_x_continuous(breaks=seq(0,10,by=0.5))+
  # scale_y_continuous(breaks=seq(0,10,by=0.5))+
  theme_bw()+
  labs(title = "Control yield - all data - no summary\nnote each site, treatment, year and rep is matched to control
       \nFacet wrap is sites",
       x = "control yield t/ha", y = "trial yield t/ha")+
  facet_wrap(.~ site_sub)

```

## Scatter Plots with Facet wrap is modification

This modification still includes everything all trials that had a modification and may amendment eg include clay fertilsier etc.

The unmodified exists because the trials might still have amendment.

Note for Jackie - the below code will need to be updated when all the sites are included and if names change.

```{r working for modification, message=FALSE, warning=FALSE, include=FALSE}
#### split the Descriptors into modification

summary_control_data_all <- summary_control_data_all %>% 
  mutate(Descriptors_temp = Descriptors) %>% 
  separate(Descriptors_temp, c("modification_with_depth", "amendment"), sep = "_")

unique(summary_control_data_all$modification_with_depth)

summary_control_data_all <- summary_control_data_all %>% 
  mutate(modification = case_when(
    
    modification_with_depth ==  "Unmodified" ~ "Unmodified",
    modification_with_depth ==  "Spade.30" ~ "Spade",
    modification_with_depth ==  "Sweep.30" ~ "Sweep",
    modification_with_depth ==  "Delving.18" ~ "Delving",
    
    modification_with_depth ==  "Rip.30" ~ "Rip",
    modification_with_depth ==  "Rip.40" ~ "Rip",
    modification_with_depth ==  "Rip.41" ~ "Rip",
    modification_with_depth ==  "Rip.50" ~ "Rip",
    modification_with_depth ==  "Rip.50IncRip" ~ "In_Rip",
    modification_with_depth ==  "Rip.60" ~ "Rip",
    modification_with_depth ==  "Rip.30+60" ~ "Rip",
    modification_with_depth ==  "Rip.60Spade.30" ~ "Rip+Spade",
    modification_with_depth ==  "Rip.30IncRip" ~ "In_Rip",
    modification_with_depth ==  "Rip.35" ~ "Rip",
    modification_with_depth ==  "Rip.40IncRip" ~ "In_Rip",
    modification_with_depth ==  "Rip.45" ~ "Rip",
    modification_with_depth ==  "Rip.45IncRip" ~ "In_Rip",
    modification_with_depth ==  "Rip.45IncRip+Spade.30" ~ "In_Rip+Spade",
    modification_with_depth ==  "Rip.60IncRip" ~ "In_Rip",
    modification_with_depth ==  "Rip.60IncRip+Spade.30" ~ "In_Rip+Spade",
    modification_with_depth ==  "Inc.50" ~ "Inc",
    modification_with_depth ==  "Delving.18" ~ "Delving",
    TRUE ~ "check"))
    

```


How many soil modification (with NO depth differences eg all the ripping is included in rip)?

```{r how many NO depth modification all, echo=FALSE, message=FALSE, warning=FALSE}
modification_NO_depth <- summary_control_data_all %>% 
  distinct(modification) %>% 
  arrange(modification)
modification_NO_depth %>% 
  count()
  
```  
  

```{r scatter modification, echo=FALSE, message=FALSE, warning=FALSE}
# ggplot(summary_control_data_all, mapping = aes(control_yield, yield,)) +
#   geom_point() +
#   geom_smooth(method = lm, se = FALSE) +
#   geom_abline(intercept = 0, slope = 1, linetype="dashed")+
#   scale_x_continuous(breaks=seq(0,10,by=0.5))+
#   scale_y_continuous(breaks=seq(0,10,by=0.5))+
#   theme_bw()+
#   labs(title = "Control yield - all data - no summary\nnote each site, treatment, year and rep is matched to control
#        \nFacet wrap is modification",
#        x = "control yield t/ha", y = "trial yield t/ha")+
#   facet_wrap(.~ modification)


summary_control_data_all %>%  filter(modification != "Unmodified") %>% 
ggplot( mapping = aes(control_yield, yield,)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  #scale_x_continuous(breaks=seq(0,10,by=0.5))+
  #scale_y_continuous(breaks=seq(0,10,by=0.5))+
  theme_bw()+
  labs(#title = "Control yield - all data - no summary\nnote each site, treatment, year and rep is matched to control
       #\nFacet wrap is modification",
       x = "control yield t/ha", y = "trial yield t/ha")+
  facet_wrap(.~ modification)


```


## subset data so there is no amendment

```{r subset data, message=FALSE, warning=FALSE, include=FALSE}
### subset data so we have no amendment
str(summary_control_data_all)
unique(summary_control_data_all$amendment)
no_amendment <- summary_control_data_all %>% 
  filter(amendment ==  "none"  )
str(no_amendment)

## change the order of the modification
oder_modification <- c("Rip","Rip+Spade", "Spade", "Sweep" , "Inc" ,"In_Rip" ,"In_Rip+Spade","Delving","Unmodified" )
no_amendment$modification <- factor(no_amendment$modification,
                                       levels = oder_modification)
```

## Scatter Plots with subset data

```{r scatter subset data, echo=FALSE, message=FALSE, warning=FALSE}
### control vs trials
ggplot(data = no_amendment, mapping = aes(control_yield, yield)) +
  geom_point(aes(colour= modification),alpha= 0.4) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", size = 0.5)+
  geom_smooth(data = no_amendment, method = lm, se = FALSE, colour = "black", size = 0.5) +
  scale_x_continuous(breaks=seq(0,10,by=0.5))+
  scale_y_continuous(breaks=seq(0,10,by=0.5))+
  theme_bw()+
  labs(title = "Control yield - subset of data no amendment",
       subtitle = "note each site, treatment, year and rep is matched to control",
       x = "control yield t/ha", y = "trial yield t/ha")



```




## Scatter Plots with subset data
### Facet wrap is years post amelioration, the NA related to trial with no metdata
```{r scatter subset data facet yrs post, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = no_amendment, mapping = aes(control_yield, yield)) +
  geom_point(aes(colour= modification),alpha= 0.4) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", size = 0.5)+
  #geom_smooth(data = no_amendment, method = lm, se = FALSE, colour = "black", size = 0.5) +
  labs(title = "Control yield - subset of data no amendment",
       subtitle = "note each site, treatment, year and rep is matched to control",
       x = "control yield t/ha", y = "trial yield t/ha")+
  facet_wrap(.~ as.factor(yr_post_amelioration))

```





## Scatter Plots with subset data
### Facet wrap is sites

```{r scatter subset data facet site, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = no_amendment, mapping = aes(control_yield, yield)) +
  geom_point(aes(colour= modification),alpha= 0.4) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", size = 0.5)+
  geom_smooth(data = no_amendment, method = lm, se = FALSE, colour = "black", size = 0.5) +
  labs(title = "Control yield - subset of data no amendment",
       subtitle = "note each site, treatment, year and rep is matched to control",
       x = "control yield t/ha", y = "trial yield t/ha")+
  facet_wrap(.~ site_sub)
```

## Scatter Plots with subset data
### Facet wrap is modification


```{r scatter subset data facet modification, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = no_amendment, mapping = aes(control_yield, yield)) +
  geom_point(aes(colour= modification),alpha= 0.4) +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", size = 0.5)+
  geom_smooth(data = no_amendment, method = lm, se = FALSE, colour = "black", size = 0.5) +
  labs(title = "Control yield - subset of data no amendment",
       subtitle = "note each site, treatment, year and rep is matched to control",
       x = "control yield t/ha", y = "trial yield t/ha")+
  facet_wrap(.~ modification)
```



# Yield Gains
## Histograms Plots


```{r yld gains cals, message=FALSE, warning=FALSE, include=FALSE}
#### Yld gain #####

#### need to make a yield gain clm first
summary_control_data_all <- summary_control_data_all %>% 
  mutate(yield_gain = yield-control_yield, na.rm = TRUE)


## cal the mean yield gains for modification
mean_values <- summary_control_data_all %>% 
  group_by(modification) %>% 
  summarise(mod_mean = mean(yield_gain,  na.rm = TRUE))

```

```{r yld gains histograms, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(summary_control_data_all, aes(yield_gain)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)), colour = 1, fill = "white") +
  geom_vline(data = summary_control_data_all, aes(xintercept=mean(yield_gain,  na.rm = TRUE)), color="blue", linetype="dashed", size=1)+
  geom_vline(data = summary_control_data_all, aes(xintercept=0), color="black", linetype="dashed", size=1)+
  labs(title = "Yield gains - all data - no summary\nNote each site, treatment, year and rep is matched to control\nBlue line is mean yield gain",
             x = "yield gain t/ha", y = "frequency of occurance")
  
```

## Histograms Plots
### Facet wrap is modification

```{r yld gains histograms facet wraps modification, echo=FALSE, message=FALSE, warning=FALSE}
  
# Histogram for each modification with mean yield gain plotted
ggplot(summary_control_data_all, aes(x = yield_gain)) + 
  # geom_histogram(aes(y = ..density..),colour = 1, fill = "white") +
  geom_histogram(aes(y = (..count..)/sum(..count..)), group = 1, colour = 1, fill = "white") +
  geom_vline(data = mean_values, aes(xintercept=mod_mean),       color="blue", linetype="dashed", size=1)+
  geom_vline(data = summary_control_data_all, aes(xintercept=0), color="black", linetype="dashed", size=1)+
  facet_wrap(.~ modification)+
  labs(title = "Yield gains - all data - no summary\nNote each site, treatment, year and rep is matched to control
         \nFacet wrap is modification\nBlue line is mean yield gain",
       x = "yield gain t/ha", y = "frequency of occurance")
```



## Histograms Plots with subset of data 
### Subset data so there is no amendment


```{r yld gains cals subset, echo=FALSE, message=FALSE, warning=FALSE}

#### need to make a yield gain clm first
no_amendment <- no_amendment %>% 
  mutate(yield_gain = yield-control_yield, na.rm = TRUE)

## cal the mean yield gains for modification
mean_values_no_amendment <- no_amendment %>% 
  group_by(modification) %>% 
  summarise(mod_mean = mean(yield_gain,  na.rm = TRUE))

```

```{r yld gains histograms subset, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(no_amendment, aes(yield_gain)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)), colour = 1, fill = "white") +
  geom_vline(data = no_amendment, aes(xintercept=mean(yield_gain,  na.rm = TRUE)), color="blue", linetype="dashed", size=1)+
  geom_vline(data = no_amendment, aes(xintercept=0), color="black", linetype="dashed", size=1)+
  labs(title = "Yield gains - subset of data no amendment",
  subtitle = "Note each site, treatment, year and rep is matched to control\nBlue line is mean yield gain",
       x = "yield gain t/ha", y = "frequency of occurance")
```

## Histograms Plots with subset of data
### Facet wrap is modification



```{r yld gains histograms subset facet wraps modification, echo=FALSE, message=FALSE, warning=FALSE}

# Histogram for each modification with mean yield gain plotted
ggplot(no_amendment, aes(yield_gain)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)), colour = 1, fill = "white") +
  geom_vline(data = mean_values_no_amendment, aes(xintercept=mod_mean), color="blue", linetype="dashed", size=1)+
  geom_vline(data = no_amendment, aes(xintercept=0), color="black", linetype="dashed", size=1)+
  labs(title = "Yield gains - subset of data no amendment",
       subtitle = "Note each site, treatment, year and rep is matched to control\nBlue line is mean yield gain",
       x = "yield gain t/ha", y = "frequency of occurance")+
  facet_wrap(.~ modification)
```


#  Cummulative yields   


```{r cum yld cals, message=FALSE, warning=FALSE, include=FALSE}

#keep the rep but not the year
cum_yld_control <- summary_control_data_all %>% 
  group_by(site_sub, Descriptors,rep_block ) %>% 
  summarise(sum_yld = sum(yield, na.rm = TRUE),
            sum__control_yld = sum(control_yield, na.rm = TRUE),
            max_yr_post_amelioration = max(yr_post_amelioration  , na.rm = TRUE))


cum_yld_control <- cum_yld_control %>% 
  mutate(`length of trial` = as.factor(max_yr_post_amelioration))
```

### Scatter plots

```{r cum yld scatter, echo=FALSE, message=FALSE, warning=FALSE}
scatter_control_vs_trial_cum_yld <- 
  ggplot(data = cum_yld_control, mapping = aes(sum__control_yld, sum_yld, )) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Cumulative control yield - all data - no summary\nnote that for each site, treatment and rep all yield is summed over all years\nand it is matched to summed control
      \nBlack line is 1:1, blue is regression line",
       x = "cumulative control yield t/ha", y = "cumulative trial yield t/ha")+
  geom_abline(intercept = 0, slope = 1, linetype="dashed")

scatter_control_vs_trial_cum_yld
```

### Scatter plots with trial lenghth

```{r cum yld scatter with lenghth , echo=FALSE, message=FALSE, warning=FALSE}
scatter_control_vs_trial_cum_yld_trial_lenghth <- 
  ggplot(data = cum_yld_control, mapping = aes(sum__control_yld, sum_yld, colour = `length of trial`)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  labs(title = "Cumulative control yield - all data - no summary",
       subtitle ="note that for each site, treatment and rep yields are summed over multiple years\nand it is matched to summed control yield",
       x = "cumulative control yield t/ha", y = "cumulative trial yield t/ha")
scatter_control_vs_trial_cum_yld_trial_lenghth
```




# Cummulative yields for yeild gains  

```{r cum yld gains cal, message=FALSE, warning=FALSE, include=FALSE}
str(summary_control_data_all)
#keep the rep but not the year
cum_yld_gains_control <- summary_control_data_all %>% 
  group_by(site_sub, Descriptors,rep_block ) %>% 
  summarise(sum_yld_gains = sum(yield_gain, na.rm = TRUE),
            max_yr_post_amelioration = max(yr_post_amelioration  , na.rm = TRUE))
cum_yld_gains_control




#### split the Descriptors into modification

cum_yld_gains_control <- cum_yld_gains_control %>% 
  mutate(Descriptors_temp = Descriptors) %>% 
  separate(Descriptors_temp, c("modification_with_depth", "amendment"), sep = "_")

unique(cum_yld_gains_control$modification_with_depth)


cum_yld_gains_control <- cum_yld_gains_control %>% 
  mutate(modification = case_when(
    
    modification_with_depth ==  "Unmodified" ~ "Unmodified",
    modification_with_depth ==  "Spade.30" ~ "Spade",
    modification_with_depth ==  "Sweep.30" ~ "Sweep",
    modification_with_depth ==  "Delving.18" ~ "Delving",
    
    modification_with_depth ==  "Rip.30" ~ "Rip",
    modification_with_depth ==  "Rip.40" ~ "Rip",
    modification_with_depth ==  "Rip.41" ~ "Rip",
    modification_with_depth ==  "Rip.50" ~ "Rip",
    modification_with_depth ==  "Rip.50IncRip" ~ "Rip",
    modification_with_depth ==  "Rip.60" ~ "Rip",
    modification_with_depth ==  "Rip.30+60" ~ "Rip",
    modification_with_depth ==  "Rip.60Spade.30" ~ "Rip",
    
    modification_with_depth ==  "Inc.50" ~ "Inc",
    modification_with_depth ==  "Delving.18" ~ "Delving",
    TRUE ~ "check"))



## cal the mean yield gains for modification
mean_gains_values <- cum_yld_gains_control %>% 
  group_by(modification) %>% 
  summarise(mod_mean_gains = mean(sum_yld_gains,  na.rm = TRUE))

cum_yld_gains_control <- cum_yld_gains_control %>% 
  mutate(`length of trial` = as.factor(max_yr_post_amelioration))
```

### Histograms plot

```{r cum yld gains histogram 1, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(cum_yld_gains_control, aes(sum_yld_gains)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)), colour = 1, fill = "white") +
  geom_vline(data = cum_yld_gains_control, aes(xintercept=mean(sum_yld_gains,  na.rm = TRUE)), color="blue", linetype="dashed", size=1)+
  geom_vline(data = cum_yld_gains_control, aes(xintercept=0), color="black", linetype="dashed", size=1)+
  labs(title = "Cumulative control yield - all data - no summary",
       subtitle ="note that for each site, treatment and rep yields are summed over multiple years\nand it is matched to summed control yield",
       x = "cumulative yield gain t/ha", y = "frequency of occurance")

```


### Histograms plot for spading and ripping only


```{r cum yld gains histogram, echo=FALSE, message=FALSE, warning=FALSE}

# Histogram for each modification with mean yield gain plotted


cum_yld_gains_control %>% 
  filter(modification =="Spade"| modification == "Rip") %>% 
 ggplot( aes(x = sum_yld_gains)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)), group = 1, colour = 1, fill = "white") +
  geom_vline(data = mean_gains_values %>% 
               filter(modification =="Spade"| modification == "Rip"), aes(xintercept=mod_mean_gains),       color="blue", linetype="dashed", size=1)+
  geom_vline(data = cum_yld_gains_control%>% 
               filter(modification =="Spade"| modification == "Rip"), aes(xintercept=0), color="black", linetype="dashed", size=1)+
  facet_wrap(.~ modification)+
  labs(title = "Cumulative control yield",
       subtitle ="note that for each site, treatment and rep yields are summed over multiple years\nand it is matched to summed control yield",
       x = "cumulative yield gain t/ha", y = "frequency of occurance")


```

# Cummulative yields for yeild gains with no amedments 


```{r cum yld gains no amedments cals, message=FALSE, warning=FALSE, include=FALSE}
#keep the rep but not the year
cum_yld_no_amendment <- no_amendment %>% 
  group_by(site_sub, Descriptors,modification,rep_block ) %>% 
  summarise(sum_yld = sum(yield, na.rm = TRUE),
            sum__control_yld = sum(control_yield, na.rm = TRUE),
            max_yr_post_amelioration = max(yr_post_amelioration  , na.rm = TRUE))
cum_yld_no_amendment

cum_yld_no_amendment <- cum_yld_no_amendment %>% 
  mutate(`length of trial` = as.factor(max_yr_post_amelioration))

```

### Scatter plots

```{r cum yld gains no amedments scatter, message=FALSE, warning=FALSE, include=FALSE}
  ggplot(data = cum_yld_no_amendment, mapping = aes(sum__control_yld, sum_yld, )) +
    geom_point(aes(colour= modification),alpha= 0.4) +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", size = 0.5)+
    geom_smooth(data = cum_yld_no_amendment, method = lm, se = FALSE, colour = "black", size = 0.5) +
      labs(title = "Cumulative control yield - subset of data no amendment",
         subtitle = "note that for each site, treatment and rep all yield is summed over all years\nand it is matched to summed control
         \nBlack dashed is 1:1, solid line is regression line",
         x = "control yield t/ha", y = "trial yield t/ha")+
  geom_abline(intercept = 0, slope = 1, linetype="dashed")

```

### Histograms

```{r cum yld gains no amedments cal mean, message=FALSE, warning=FALSE, include=FALSE}
  
  ## cal the mean yield gains for modification
  mean_gains_values_no_amendment <- cum_yld_no_amendment %>% 
    group_by(modification) %>% 
    summarise(mod_mean_gains = mean(sum_yld,  na.rm = TRUE))
  
``` 
  
 
```{r cum yld gains no amedments hist, echo=FALSE, message=FALSE, warning=FALSE}
 
  cum_yld_no_amendment %>% 
    ggplot( aes(x = sum_yld)) + 
    geom_histogram(aes(y = (..count..)/sum(..count..)), group = 1, colour = 1, fill = "white") +
    
    geom_vline(data = mean_gains_values_no_amendment, 
               aes(xintercept=mod_mean_gains),       color="blue", linetype="dashed", size=1)+
    
    geom_vline(data = cum_yld_no_amendment, 
               aes(xintercept=0), color="black", linetype="dashed", size=1)+
    facet_wrap(.~ modification)+
    labs(title = "Cumulative control yield - subset of data no amendment",
         subtitle ="note that for each site, treatment and rep yields are summed over multiple years\nand it is matched to summed control yield",
         x = "cumulative yield gain t/ha", y = "frequency of occurance")
  
```


### Histograms just the ripping and spading
```{r cum yld gains no amedments hist spad and rip, echo=FALSE, message=FALSE, warning=FALSE}
  cum_yld_no_amendment %>% 
    filter(modification =="Spade"| modification == "Rip") %>% 
    ggplot( aes(x = sum_yld)) + 
    geom_histogram(aes(y = (..count..)/sum(..count..)), group = 1, colour = 1, fill = "white") +
    
    geom_vline(data = mean_gains_values_no_amendment%>% 
                filter(modification =="Spade"| modification == "Rip"), 
               aes(xintercept=mod_mean_gains),       color="blue", linetype="dashed", size=1)+
    
    geom_vline(data = cum_yld_no_amendment %>% 
                filter(modification =="Spade"| modification == "Rip"), 
               aes(xintercept=0), color="black", linetype="dashed", size=1)+
    facet_wrap(.~ modification)+
    labs(title = "Cumulative control yield - subset of data no amendment",
         subtitle ="note that for each site, treatment and rep yields are summed over multiple years\nand it is matched to summed control yield",
         x = "cumulative yield gain t/ha", y = "frequency of occurance")
```
